The semantic drift of quotations in blogspace: a case study in short-term cultural evolution

### 2016-08-23 brains-introduction.tex
okey-dokey manuscrit

We thus ask the following question: given such representations that seem to evolve precisely because of the kind of automatic cognitive biases evoked in the theory of epidemiology of representations, \tb{do cultural attractors appear, and if so how do cognitive biases participate in them?}
%how do cognitive biases participate in \tb{the appearance of cultural attractors}?
We chose to restrict our analysis to substitutions (i.e., one word being replaced by another), both to keep the analysis tractable and because of missing information in our data set.\footnote{
\newtext{As explained further down, source-destination links between quotes must be inferred from the data set, an operation which is much more reliable if we restrict our analysis to substitutions.
This also impedes us from observing the effect of accumulated transformations in the long term, limiting our results to a view of the individual evolutionary step.
}}
While this limits the scope of our results to the particular data set we use, the methodological point we also make is left intact.  By characterizing words using 6 well-studied features, we identify what makes a substitution more likely, and how a word changes when it is substituted.
This exploratory approach uncovers a number of transmission biases consistent with known effects in linguistics.
While the transformations we describe are not the only ones at work in this data set, our analysis also indicates that feature-specific attractors could exist because of the substitution process.
This study can be viewed as analyzing part of the transmission step operating in transmission chains of artificial languages like those studied by \citet{kirby_cumulative_2008}, yet with natural language out of the laboratory.

>>>>

We thus ask the following question: given such representations that seem to evolve precisely because of the kind of automatic cognitive biases evoked in the theory of epidemiology of representations, do cultural attractors appear, and if so how do cognitive biases participate in them?
%how do cognitive biases participate in \tb{the appearance of cultural attractors}?
We chose to restrict our analysis to substitutions (i.e., one word being replaced by another), both to keep the analysis tractable and because of missing information in our data set.\footnote{
\newtext{As explained further down, source-destination links between quotes must be inferred from the data set, an operation which is much more reliable if we restrict our analysis to substitutions.
This also impedes us from observing the effect of accumulated transformations in the long term, limiting our results to a view of the individual evolutionary step.
}}
While this limits the scope of our results to the particular data set we use, the methodological point we also make is left intact.  By characterizing words using 6 well-studied features, we identify what makes a substitution more likely, and how a word changes when it is substituted.
This exploratory approach uncovers a number of transmission biases consistent with known effects in linguistics.
While the transformations we describe are not the only ones at work in this data set, our analysis also indicates that feature-specific attractors could exist because of the substitution process.
This study can be viewed as analyzing part of the transmission step operating in transmission chains of artificial languages like those studied by \citet{kirby_cumulative_2008}, yet with natural language out of the laboratory.

### 2016-08-23 brains-protocol.tex
Move figures around to bring the closer to their citing text
okey-dokey manuscrit

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{images/computed-figures/all-susceptibilities-pos.png}
    \caption{\newtext{\textbf{POS-related results:}
    categories are simplified from the TreeTagger tag set: \tb{\emph{C} means \emph{Closed class-like}} (see main text for details), \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, and \emph{V} means verb.
    The top panel shows the actual $s_{POS}$ and $s_{POS}^0$ counts. The bottom panel shows the substitution susceptibility $\sigma_{POS}$, which is the ratio between the two previous counts.
    Confidence intervals are computed with the \citet{goodman_simultaneous_1965} method for multinomial proportions.}
    }
    \label{fig:pos-susceptibilities}
\end{figure}

@@

@@

Fig.~\ref{fig:pos-susceptibilities} gathers the results for POS groups.
A Goodman-based multinomial goodness-of-fit test~\citep{goodman_simultaneous_1965} shows that these categories have a significant effect on susceptibility ($p < .05$ in all substitution models), but this seems mostly due to the \emph{Closed class-like}\footnote{
\newtext{The \emph{Closed class-like} category gathers all the POS groups \tb{representing closed class words (coordinating conjunctions, prepositions, subordinating conjunctions, modals and possessive endings).
These groups, essentially made of stopwords, feature very low counts for both $s$ (substitutions falling on stopwords are filtered out) and $s^0$ (stopwords are never counted as substitutable).
While the susceptibility reported for the remaining words is left unbiased (as $s$ and $s^0$ are equally affected), they represent a very small portion of all substitutions, which led us to group them together.
Finally, we added to this meta-category the few POS groups that cover words entirely excluded from the analysis (foreign words, punctuation symbols and interjections), only sporadically present because of tagging fluctuations;
hence the name \emph{Closed class-like}.
}}
}
and \emph{Adverb} categories.
Indeed, detailing which categories are out of their confidence region under $\mathcal{H}_0$ shows that susceptibility for closed class-likes is significantly below 1 (confirmed in all substitution models), as is that for adverbs in 13 of the 16 substitution models;
none of the other categories are significantly different from $\mathcal{H}_0$ (except nouns which appear significantly above 1 in a single substitution model).
While we acknowledge the low susceptibilities of adverbs and closed class-likes, these categories concern less than 7\% of all substitutions under $\mathcal{H}_0$ (and even less in the actual data);
it seems, then, that POS categories do not capture any strong bias in the selection of substitution targets.

>>>>

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{images/computed-figures/all-susceptibilities-pos.png}
    \caption{\newtext{\textbf{POS-related results:}
    categories are simplified from the TreeTagger tag set: \emph{C} means \emph{Closed class-like} (see main text for details), \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, and \emph{V} means verb.
    The top panel shows the actual $s_{POS}$ and $s_{POS}^0$ counts. The bottom panel shows the substitution susceptibility $\sigma_{POS}$, which is the ratio between the two previous counts.
    Confidence intervals are computed with the \citet{goodman_simultaneous_1965} method for multinomial proportions.}
    }
    \label{fig:pos-susceptibilities}
\end{figure}

@@

Fig.~\ref{fig:pos-susceptibilities} gathers the results for POS groups.
A Goodman-based multinomial goodness-of-fit test~\citep{goodman_simultaneous_1965} shows that these categories have a significant effect on susceptibility ($p < .05$ in all substitution models), but this seems mostly due to the \emph{Closed class-like}\footnote{
\newtext{The \emph{Closed class-like} category gathers all the POS groups representing closed class words (coordinating conjunctions, prepositions, subordinating conjunctions, modals and possessive endings).
These groups, essentially made of stopwords, feature very low counts for both $s$ (substitutions falling on stopwords are filtered out) and $s^0$ (stopwords are never counted as substitutable).
While the susceptibility reported for the remaining words is left unbiased (as $s$ and $s^0$ are equally affected), they represent a very small portion of all substitutions, which led us to group them together.
Finally, we added to this meta-category the few POS groups that cover words entirely excluded from the analysis (foreign words, punctuation symbols and interjections), only sporadically present because of tagging fluctuations;
hence the name \emph{Closed class-like}.
}
}
and \emph{Adverb} categories.
Indeed, detailing which categories are out of their confidence region under $\mathcal{H}_0$ shows that susceptibility for closed class-likes is significantly below 1 (confirmed in all substitution models), as is that for adverbs in 13 of the 16 substitution models;
none of the other categories are significantly different from $\mathcal{H}_0$ (except nouns which appear significantly above 1 in a single substitution model).
While we acknowledge the low susceptibilities of adverbs and closed class-likes, these categories concern less than 7\% of all substitutions under $\mathcal{H}_0$ (and even less in the actual data);
it seems, then, that POS categories do not capture any strong bias in the selection of substitution targets.
